---
title: "Sero and cases measles analysis, Children Hospital 2"
title-block-banner: "#002147"
format:
  html:
    toc: true
    toc-location: left
    toc-title: Summary
    code-fold: true
    number-sections: true
editor: visual
message: false
warning: false
echo: true
bibliography: references.bib
---

# Setup

```{r}
#| code-summary: "Library"

library(readxl)
library(tidyverse)
library(lubridate)
library(sf)
library(janitor)
library(stringr)
library(magrittr)
library(stringi)
library(paletteer)
library(patchwork)
library(epitools)
invisible(Sys.setlocale("LC_TIME", "English"))
```

```{r,results='hide'}
#| code-summary: "Input data"

### measles cases 2024-2025
cases_2425 <- read_excel("D:/OUCRU/sero_measles_nd2/data/Measles2018-2019-2024-2025_17.06.25.xlsx",sheet = "2024-2025")

### measles cases 2018
cases_18 <- read_excel("D:/OUCRU/sero_measles_nd2/data/Measles2018-2019-2024-2025_17.06.25.xlsx", 
                         sheet = "2018")

### measles cases 2019
cases_19 <- read_excel("D:/OUCRU/sero_measles_nd2/data/Measles2018-2019-2024-2025_17.06.25.xlsx", 
                       sheet = "2019")

### all measles sero (ndtp, nd2)
sero <-  read_excel("D:/OUCRU/sero_measles_nd2/data/20240811_measles_titer_hcdc_with_meta.xlsx")

### vaccine registry data
vaxreg_hcmc_measles <- readRDS("D:/OUCRU/sero_measles_nd2/data/vaxreg_hcmc_measles.rds")

## hcmc population
census2019 <- readRDS("D:/OUCRU/hfmd/data/census2019.rds") %>% 
  filter(province == "Thành phố Hồ Chí Minh") %>% 
  mutate(age2 = word(age,1),
         age2 = as.numeric(age2))

hcm_pop_19 <- census2019 %>% mutate(
  district = district %>% 
    str_replace_all(
      c("Quận 2" = "Thủ Đức",
        "Quận 9" = "Thủ Đức")) %>% 
    str_remove("Quận|Huyện") %>%
    trimws(which = "both") %>% 
    stri_trans_general("latin-ascii") %>% 
    tolower()) %>% 
    filter(age2 <= 15) %>% 
    group_by(district) %>% 
    summarise(pop = sum(n))


### HCMC map
map_path <- "D:/OUCRU/HCDC/project phân tích sero quận huyện/"
vn_qh <- st_read(dsn = file.path(map_path, "gadm41_VNM.gpkg"), layer = "ADM_ADM_2")

vn_qh1 <- vn_qh %>%
  clean_names() %>%     ## cho thành chữ thường
  filter(
    str_detect(
      name_1,
      "Hồ Chí Minh"
    )
  )
qhtp <- vn_qh1[-c(14,21),]

qhtp$geom[qhtp$varname_2 == "Thu Duc"] <- vn_qh1[c("21","24","14"),] %>%
  st_union()

qhtp <- qhtp %>% st_cast("MULTIPOLYGON")

qhtp$varname_2 <- stri_trans_general(qhtp$varname_2, "latin-ascii") %>%
  tolower() %>%
  str_remove("district") %>%
  trimws(which = "both")

qhtp$nl_name_2 <- c("BC","BTân","BT","CG","CC","GV",
                    "HM","NB","PN","1","10","11","12",
                    "3","4","5","6","7","8","TB",
                    "TP","TĐ")

```

# Data description

## Cases

```{r}
#| code-summary: "2024 Cases analysis"
#| fig-width: 8
#| fig-height: 6 
#| out-width: "100%"
incidence_hcm <- cases_2425 %>% 
  filter(province == "TP.HCM") %>%
  distinct(.keep_all = TRUE)

incidence_hcm <- incidence_hcm %>% mutate(
  district2 = district %>% 
    str_replace_all(
      c("Quận 2" = "Thủ Đức",
        "Quận 9" = "Thủ Đức")) %>% 
    str_remove("Quận|Huyện") %>%
    trimws(which = "both") %>% 
    stri_trans_general("latin-ascii") %>% 
    tolower() ,
  agegr2 = factor(agegr,
                  levels = c("< 6th","6 - 9th","9th -12th","1 - 5T","6-10T","> 10T"),
                  labels = c("< 6m","6-9m","9-12m","1-5y","6-10y","> 10y")),
  adm_month = month(admission),
  adm_year = year(admission),
  adm_month_cut = cut(adm_month,
                      breaks = c(0.5,3,6,9,12),
                      labels = c("Jan-Mar","Apr-June","Jul-Sep","Oct-Dec")),
  adm_month_year_cut = str_c(adm_month_cut,adm_year,sep = " ") %>% 
    factor(levels = c("Apr-June 2024","Jul-Sep 2024","Oct-Dec 2024",
                      "Jan-Mar 2025","Apr-June 2025"))
) 

tdnd2 <- data.frame(long = 106.7025,
                    lat  = 10.7808) %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326)


incidene_36912 <- incidence_hcm %>% 
  group_by(adm_month_year_cut,district2) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(district2) %>% 
  left_join(.,hcm_pop_19, by = join_by(district2 == district)) %>% 
  mutate(prevalence_10000 = (n/pop)*10000) %>% 
  ungroup() %>% 
  group_by(adm_month_year_cut) %>% 
  group_modify(~.x %>% mutate(prevalence = n/sum(n)) %>% 
                 left_join(qhtp, ., by = join_by(varname_2 == district2))) %>% 
  ggplot() +
  geom_sf(aes(fill = prevalence_10000,geometry = geom),show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Prevalence per 10.000 children")+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  facet_wrap(~factor(adm_month_year_cut,
                     levels = c("Apr-June 2024","Jul-Sep 2024","Oct-Dec 2024",
                                "Jan-Mar 2025","Apr-June 2025")),ncol = 3) +
  labs(title = "Prevalence in 2024-2025 outbreak")+
  theme_void()+
  theme(legend.position = "bottom",
        legend.key.width =  unit(1, "cm"))

incidene_36912
```

```{r}
#| code-summary: "2018-2019 Cases analysis"
#| fig-width: 8
#| fig-height: 6 
#| out-width: "100%"
cases_hcm_18 <- cases_18 %>% filter(province == "TP. H.C.M") %>%
  distinct(.keep_all = TRUE) %>% 
  select(id,gender,dob,agegr,district,admission)

cases_hcm_19 <- cases_19 %>% filter(province == "TP. H.C.M")%>%
  distinct(.keep_all = TRUE) %>% 
  select(id,gender,dob,agegr,district,admission)

cases_hcm_1819 <- rbind(cases_hcm_18,cases_hcm_19) %>% 
  distinct(id,dob, .keep_all = TRUE)  %>% mutate(
    district2 = district %>% 
      str_replace_all(
        c("Quận 2" = "Thủ Đức",
          "Quận 9" = "Thủ Đức")) %>% 
      str_remove("Quận|Huyện") %>%
      trimws(which = "both") %>% 
      stri_trans_general("latin-ascii") %>% 
      tolower() ,
    agegr2 = factor(agegr,
                    levels = c("<6 THANG","6-9 THANG","9Thang -1 Tuoi",
                               "Từ 1-5 tuổi","Từ 5-10 tuổi","Trên 10 tuổi"),
                    labels = c("< 6m","6-9m","9-12m","1-5y","6-10y","> 10y")),
    adm_month = month(admission),
    adm_year = year(admission),
    adm_month_cut = cut(adm_month,
                        breaks = c(0.5,3,6,9,12),
                        labels = c("Jan-Mar","Apr-June","Jul-Sep","Oct-Dec")),
    adm_month_year_cut = str_c(adm_month_cut,adm_year,sep = " ")
  ) 


cases_hcm_1819 %>% 
  group_by(adm_month_year_cut,district2) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(district2) %>% 
  left_join(.,hcm_pop_19, by = join_by(district2 == district)) %>% 
  mutate(prevalence_10000 = (n/pop)*10000) %>% 
  ungroup() %>% 
  group_by(adm_month_year_cut) %>% 
  group_modify(~.x %>% mutate(prevalence = n/sum(n)) %>% 
                 left_join(qhtp, ., by = join_by(varname_2 == district2))) %>% 
  ggplot() +
  geom_sf(aes(fill = prevalence_10000,geometry = geom),show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Prevalence per 10.000 children")+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  facet_wrap(~factor(adm_month_year_cut,
                     levels = c("Apr-June 2018","Jul-Sep 2018","Oct-Dec 2018",
                                "Jan-Mar 2019","Apr-June 2019","Jul-Sep 2019",
                                "Oct-Dec 2019")),
             ncol = 4) +
  labs(title = "Prevalence in 2018-2019 outbreak")+
  theme_void()+
  theme(legend.position = "bottom",
        legend.key.width =  unit(1, "cm"))
```

## Sero

```{r}
#| code-summary: "Sero analysis"
#| fig-width: 8
#| fig-height: 6 
#| out-width: "100%"

sero_nd2 <- sero %>% filter(hospital == "Bv Nhi Dong 2") %>% 
  select(pos,district,age,age_1y,age_5y,sampling_period,dob) %>% 
  mutate(
    district2 = district %>% 
      stri_trans_general("latin-ascii") %>% 
      str_remove("Tp|^0") %>%
      trimws(which = "both") %>% 
      tolower() ,
    sampling_period = as.Date(sampling_period),
    dob = as.Date(dob),
    samp_month = month(sampling_period),
    samp_year = year(sampling_period)
    )

sero_negative_district <- sero_nd2 %>% 
  group_by(sampling_period,district2) %>% 
  count(pos) %>% 
  pivot_wider(names_from = pos, 
              values_from = n,
              names_prefix = "pos_") %>% 
  replace(is.na(.), 0) %>% 
  ungroup() %>% 
  mutate(total = pos_0+pos_1,
         sp = pos_1/total,
         sneg = 1 - sp)  %>% 
  group_by(sampling_period) %>% 
  group_modify(~.x %>% left_join(qhtp, ., by = join_by(varname_2 == district2))) %>% 
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = sneg,geometry = geom),show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    labels = scales::label_percent(),
                                    na.value="white",
                                    name = "Sero-negative")+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  facet_wrap(~factor(sampling_period,
                     labels = c("Sep 2022","Dec 2022","Mar 2023",
                                "June 2023","Oct 2023","Dec 2023")),
             ncol = 3) +
  labs(title = "Sero-negative from Sep 2022 - Dec 2023")+
  theme_void()+
  theme(legend.position = "bottom")
  
sero_negative_district  
```

## Total number of cases and sero samples

### Children's Hospital 2 {.unnumbered}

```{r}
#| fig-width: 8
#| fig-height: 6 
#| out-width: "100%"

p_num_cases_24 <- incidence_hcm %>% 
  group_by(district2) %>% 
  count() %>% 
  left_join(qhtp, ., by = join_by(varname_2 == district2)) %>%
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = n,geometry = geom),show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    # labels = scales::label_percent(),
                                    na.value="white",
                                    name = "Number of cases")+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  labs(title = "2024-2025")+
  theme_void()+
  theme(legend.position = "bottom")

p_num_cases_18 <- cases_hcm_1819 %>% 
  group_by(district2) %>% 
  count() %>% 
  left_join(qhtp, ., by = join_by(varname_2 == district2)) %>%
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = n,geometry = geom),show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    # labels = scales::label_percent(),
                                    na.value="white",
                                    name = "Number of cases")+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  labs(title = "2018-2019")+
  theme_void()+
  theme(legend.position = "bottom")

p_num_sero <-sero_nd2 %>% 
  group_by(district2) %>% 
  count() %>% 
  left_join(qhtp, ., by = join_by(varname_2 == district2)) %>%
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = n,geometry = geom),show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Number of \n serum samples")+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  labs(title = "Dec 2022 - Dec 2023")+
  theme_void()+
  theme(legend.position = "bottom")

p_num_cases_18 | p_num_cases_24 | p_num_sero

```

## Correlation between number of cases and serum samples

```{r}
#| fig-width: 8
#| fig-height: 5 
#| out-width: "100%"

library(ggstatsplot)

num_sero <- sero_nd2 %>% 
  group_by(district2) %>% 
  count()

numcase_24 <- incidence_hcm %>% 
  group_by(district2) %>% 
  count()


left_join(numcase_24,num_sero,by = join_by(district2)) %>% 
  set_colnames(c("district","num_cases","num_sero")) %>% 
  ggscatterstats(
    x = num_cases,
    y = num_sero,
    bf.message = FALSE,
    marginal = FALSE, 
    label.var = district,
    xlab = "Number of measles cases in 2024",
    ylab = "Number of serum samples from 2022"
    )
```

# Catchment area

Catchment area of Children Hospital 2 was calculated by cummulative cases ratio method [@zinszer2014].

The cumulative case ratio was defined as the ratio of the observed to the expected number of measles visits to CH2 from a district. The expected number of measles visits to CH2 from a particular district was calculated by multiplying the district’s population by the cumulative case rate for that hospital.

$$
\begin{align}
Cummulative \ cases \ rate &= \frac{observed \ cases \ (districts)}{expected \ cases \ (districts)} \\ &= \frac{observed \ cases \ (districts)}{district \ population \times cummulative \ cases \  rate \ (hospitals)}
\end{align}
$$

$$
cummulative \ cases \  rate \ (hospitals)= \frac{cases \ observed \ (hospital)}{population \ of \ children \ under \ 15 \ in \ HCMC}
$$

::: callout-note
A district was included in the catchment area if the upper limit of the 95% confidence interval for the cumulative case ratio for that district was 1 or greater because a ratio less than 1 indicated that the district contributed significantly fewer measles visits than expected for its population
:::

::: {.panel-tabset .nav-pills}
## 2024-2025

```{r}
#| code-summary: "Catchment area"
#| fig-width: 8
#| fig-height: 6 
#| out-width: "100%"



ch2_obs <- incidence_hcm %>% nrow()

ch2_cum_case_rate <- ch2_obs/sum(hcm_pop_19$pop)


catchment_ch2_2425 <- incidence_hcm %>% 
  group_by(district2) %>% 
  count() %>% 
  left_join(hcm_pop_19,., by = join_by(district == district2)) %>% 
  replace(is.na(.), 0) %>% 
  mutate(expected_cases = pop*ch2_cum_case_rate,
         cum_case_ratio = n/expected_cases,
         lwr = pois.exact(n)$lower/expected_cases,
         upr = pois.exact(n)$upper/expected_cases,
         cut = cut(upr,
                   breaks = c(0,1,2,3,10),
                   labels = c("< 1",">= 1",">= 2",">= 3"),
                   right = F)
         ) 

catchment_ch2_2425 %>% 
  left_join(qhtp, ., by = join_by(varname_2 == district))%>% 
  ggplot() +
  geom_sf(aes(fill = cut,geometry = geom),show.legend = T)+
  scale_fill_manual(
    values = c(
      "< 1" = "#FFFFFFFF",
      ">= 1" = "#6BAED6FF",
      ">= 2" = "#2171B5FF",
      ">= 3" = "#08306BFF"
               ),
    name = "95% CI Upper bound of \n cumulative case ratio of each districts"
  )+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=2,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  theme_void()
```

## 2018-2019

```{r}
#| code-summary: "Catchment area"
#| fig-width: 8
#| fig-height: 6 
#| out-width: "100%"

ch2_obs_1819 <- cases_hcm_1819 %>% nrow()

ch2_cum_case_rate <- ch2_obs_1819/sum(hcm_pop_19$pop)

catchment_ch2_1819 <- cases_hcm_1819 %>% 
  group_by(district2) %>% 
  count() %>% 
  left_join(hcm_pop_19,., by = join_by(district == district2)) %>% 
  replace(is.na(.), 0) %>% 
  mutate(expected_cases = pop*ch2_cum_case_rate,
         cum_case_ratio = n/expected_cases,
         lwr = pois.exact(n)$lower/expected_cases,
         upr = pois.exact(n)$upper/expected_cases,
         cut = cut(upr,
                   breaks = c(0,1,2,3,10),
                   labels = c("< 1",">= 1",">= 2",">= 3"),
                   right = F)
         ) 

catchment_ch2_1819 %>% 
  left_join(qhtp, ., by = join_by(varname_2 == district)) %>% 
  ggplot() +
  geom_sf(aes(fill = cut,geometry = geom),show.legend = T)+
  scale_fill_manual(
    values = c(
      "< 1" = "#FFFFFFFF",
      ">= 1" = "#6BAED6FF",
      ">= 2" = "#2171B5FF",
      ">= 3" = "#08306BFF"
               ),
    name = "95% CI Upper bound of \n cumulative case ratio of each districts"
  )+
  geom_sf_text(aes(label = nl_name_2,geometry = geom),size=2,color = "black")+
  geom_sf(data = tdnd2, shape = 17,
          color = "yellow", size = 1)+
  theme_void()
```
:::

::: callout-note
The results show a similarity in the CH2 catchment areas of the two outbreaks (2018-2019 and 2024-2025). Hoc Mon was part of the CH2 catchment area in 2018-2019 but not in 2024-2025.
:::


# Analysis data from EOC and serum from 3 hospital

## Data from HCDC

```{r}
#| fig-width: 10
#| fig-height: 8 
#| out-width: "100%"

## data description for CCH and CH1
measles_1825_eoc <- read_excel("D:/OUCRU/sero_measles_nd2/data/SOI_2018-2025_EOC.xlsx", 
                                sheet = "SOI_2018-2025")

sero_ch1 <- read_csv("D:/OUCRU/sero_measles_nd2/data/20240808_measles_titer_oucru.csv")

## filter bv nhi

# measles_1825_eoc %>% 
#   clean_names() %>% 
#   filter(grepl('Nhi|nhi', ten_bv)) %>% 
#   pull(ten_bv) %>% unique()

meases_1812_3hos <- measles_1825_eoc %>% 
  clean_names() %>% 
  filter(ten_bv %in% c("Bệnh viện Nhi đồng 1",
                       "Bệnh viện Nhi Đồng 1",
                       "BV Nhi Đồng 1",
                       "Bệnh viện Nhi đồng thành phố",
                       "Bệnh viện Nhi đồng 2",
                       "Bệnh viện Nhi Đồng 2")) %>% 
  mutate(
    district = quan_huyen %>% 
      str_replace_all(
        c("Quận 2" = "Thủ Đức",
          "Quận 9" = "Thủ Đức")) %>% 
      str_remove("Quận|Huyện|Thành phố|QUẬN") %>%
      trimws(which = "both") %>% 
      stri_trans_general("latin-ascii") %>% 
      tolower(),
    admission = as.Date(ngay_nv,format = "%d/%m/%Y"),
    hospital = ten_bv %>% 
      str_replace_all(
        c("BV Nhi Đồng 1" = "Bệnh viện Nhi Đồng 1")) %>%
      trimws(which = "both") %>% 
      stri_trans_general("latin-ascii") %>% 
      tolower() %>% 
      factor(levels = c("benh vien nhi dong 1",
                        "benh vien nhi dong 2",
                        "benh vien nhi dong thanh pho"),
             labels = c("Children's Hospital 1",
                        "Children's Hospital 2",
                        "City Children's Hospital")),
    out_year = case_when(
      nam_nv %in% c(2018,2019) ~ "2018-2019",
      nam_nv %in% c(2024,2025) ~ "2024-2025"
    ),
    dob = as.Date(ngay_sinh,format="%d/%m/%Y"),
    age2 = interval(dob, admission) / years(1),
    age_gr = cut(age2,
                 breaks = c(0,5,10,15,100),
                 labels = c("0-5","5-10","10-15",">15"))
  ) %>% 
  filter(district %in% qhtp$varname_2) %>%
  distinct(.keep_all = TRUE)


num_cases_3hos <- meases_1812_3hos %>% 
  filter(!is.na(out_year)) %>% 
  group_by(hospital,district,out_year) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(hospital,out_year) %>% 
  group_modify(~.x %>% left_join(qhtp, ., by = join_by(varname_2 == district))) %>% 
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = n,geometry = geom),
          show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Number of cases")+
  # geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  # geom_sf(data = tdnd2, shape = 17,
  #         color = "yellow", size = 1)+
  facet_grid(hospital ~ out_year,
             switch = "y") +
  theme_void()+
  theme(legend.position = "bottom",
        legend.key.width =  unit(1, "cm"))

## dec 2022 - Apr 2024
num_sero_ch1 <- sero_ch1 %>% 
  select(district,doc,moc,yoc) %>% 
  na.omit(district) %>% 
  mutate(district = district %>%
           trimws(which = "both") %>% 
           stri_trans_general("latin-ascii") %>% 
           tolower()) %>% 
  group_by(district) %>% 
  count() %>% 
  mutate(hospital = rep("nd1"),
         col = rep("Dec 2022 - Apr 2024"))%>% 
  ungroup()

num_sero_ch2_cch <- sero  %>% 
  select(pos,district,age,age_1y,age_5y,sampling_period,hospital) %>% 
  mutate(
    district = district %>% 
      stri_trans_general("latin-ascii") %>% 
      str_remove("Tp|^0") %>%
      trimws(which = "both") %>% 
      tolower() ,
    samp_month = month(sampling_period),
    samp_year = year(sampling_period)
  ) %>% 
  group_by(hospital,district) %>% 
  count() %>% 
  mutate(col = rep("Dec 2022 - Dec 2023")) %>% 
  ungroup()


num_sero_3hos_p <- rbind(num_sero_ch1,num_sero_ch2_cch) %>% 
  group_by(hospital) %>% 
  group_modify(~.x %>% left_join(qhtp, ., by = join_by(varname_2 == district))) %>% 
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = n,geometry = geom),
          show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Number of \nserum samples")+
  # geom_sf_text(aes(label = nl_name_2,geometry = geom),size=1.5,color = "black")+
  # geom_sf(data = tdnd2, shape = 17,
  #         color = "yellow", size = 1)+
  facet_wrap(~factor(hospital,
                     levels = c("nd1","Bv Nhi Dong 2","Bv Nhi Dong Tp"),
                     labels = c("Dec 2022 - Apr 2024",
                                "Dec 2022 - Dec 2023",
                                "Dec 2022-Dec 2023")),
             ncol = 1) +
  theme_void()+
  theme()
  
num_cases_3hos|num_sero_3hos_p
```

## Compare seroprevalence with and without CH2

```{r}
#| fig-width: 10
#| fig-height: 8 
#| out-width: "100%"
#| fig-cap: "Figure A shows the age-dependent seroprevalence profile for April 2024 from serum samples collected at 3 hospitals. Figure B shows the age-dependent seroprevalence profile for April 2024 from serum samples excluding those from Children’s Hospital 2."

library(mgcv)

sero_ch1a <- sero_ch1 %>% 
  select(commune,district,doc,moc,yoc,dob,mob,yob,pos,age) %>% 
  na.omit(district) %>% 
  mutate(district = district %>%
           trimws(which = "both") %>% 
           stri_trans_general("latin-ascii") %>% 
           tolower(),
         col_time = make_date(yoc,moc,doc),
         hospital = rep("CH1"),
         dob = make_date(yob,mob,dob))

sero_ch2_cch <- sero  %>% 
  select(pos,commune,district,age,age_1y,age_5y,sampling_period,hospital,doc,dob) %>% 
  mutate(
    district = district %>% 
      stri_trans_general("latin-ascii") %>% 
      str_remove("Tp|^0") %>%
      trimws(which = "both") %>% 
      tolower() ,
    samp_month = month(sampling_period),
    samp_year = year(sampling_period),
    col_time = doc
  )

sero_3bv <- rbind(sero_ch1a[,c("age","commune","district","pos","col_time","dob","hospital")],
      sero_ch2_cch[,c("age","commune","district","pos","col_time","dob","hospital")]) %>% 
  mutate(age2 = round(age)) %>% 
  # left_join(.,district_xy, by = join_by(district)) %>% 
  mutate(yoc = year(col_time),
         moc = month(col_time),
         doc = day(col_time),
         age_at_apr24 = interval(dob, as.Date("2024-04-30")) / years(1),
         age424 = round(age_at_apr24))


age_profile <- function(data, age_values = seq(0, 16, le = 512), ci = .95) {
  model <- data %>% 
    group_by(age424) %>%
    count(pos) %>% 
    pivot_wider(names_from = pos,
                values_from = n,
                names_prefix = "pos_") %>% 
    replace(is.na(.), 0) %>% 
    ungroup() %>% 
    mutate(total = pos_0+pos_1,
           sp = pos_1/total,
           sneg = 1 - sp) %>% 
    gam(cbind(pos_1,total)~s(age424,bs = "cr"),
        family = binomial(link = "cloglog"),
        method = "REML",
        data = .)
  
  link_inv <- family(model)$linkinv
  df <- nrow(data) - length(coef(model))
  p <- (1 - ci) / 2
  
  model |> 
    predict(list(age424 = age_values), se.fit = TRUE) %>%
    c(list(age = age_values), .) |> 
    as_tibble() |> 
    mutate(lwr = link_inv(fit + qt(    p, df) * se.fit),
           upr = link_inv(fit + qt(1 - p, df) * se.fit),
           fit = link_inv(fit)) |> 
    select(- se.fit)
}


age_profile_apr24 <- age_profile(sero_3bv) %>% 
  mutate(sus = 1-fit,
         lwr_s = 1-lwr,
         upr_s = 1-upr) %>%  
  ggplot(aes(x = age,y = fit))+
  geom_line()+
  geom_ribbon(aes(ymin = lwr,ymax = upr),fill = "blue",alpha = 0.4)+
  scale_y_continuous(limits = c(0,1),
                     name = "Seroprevalence (%)",
                     labels = scales::label_percent())+
  scale_x_continuous(limits = c(0,16),
                     breaks = seq(0,16,by=2),
                     name = "Age (years)")+
  labs(tag = "A")+
  theme_bw()


age_profile_apr24_wo_ch2 <- sero_3bv %>% filter(hospital != "Bv Nhi Dong 2") %>% 
age_profile() %>% 
  mutate(sus = 1-fit,
         lwr_s = 1-lwr,
         upr_s = 1-upr) %>%  
  ggplot(aes(x = age,y = fit))+
  geom_line()+
  geom_ribbon(aes(ymin = lwr,ymax = upr),fill = "blue",alpha = 0.4)+
  scale_y_continuous(limits = c(0,1),
                     name = "Seroprevalence (%)",
                     labels = scales::label_percent())+
  scale_x_continuous(limits = c(0,16),
                     breaks = seq(0,16,by=2),
                     name = "Age (years)")+
  labs(tag = "B")+
  theme_bw()

age_profile_apr24 + age_profile_apr24_wo_ch2
```

```{r}
#| fig-width: 10
#| fig-height: 8 
#| out-width: "100%"
#| fig-cap: "Age density distribution of serum samples collected from 3 hospital"

sero_3bv %>% 
  ggplot(aes(x = age,
             color = factor(hospital,
                            labels = c("CH2","CCH","CH1"))))+
  geom_density()+
  theme_minimal()+
  labs(color = "Hospitals",x = "Age (years)", y = "Density")+
  theme(legend.position = "bottom")
```

```{r}
#| fig-width: 10
#| fig-height: 8 
#| out-width: "100%"
#| fig-cap: "Age density distribution in April 2024 of serum samples collected from 3 hospitals"

sero_3bv %>% 
  ggplot(aes(x = age_at_apr24,
             color = factor(hospital,
                            labels = c("CH2","CCH","CH1"))))+
  geom_density()+
  theme_minimal()+
  scale_x_continuous(breaks = seq(0,16,by = 2))+
  labs(color = "Hospitals",x = "Age in Apr 2024 (years)", y = "Density")+
  theme(legend.position = "bottom")
```



```{r}
#| fig-width: 10
#| fig-height: 8 
#| out-width: "100%"
#| fig-cap: "Figure A shows the spatial distribution of serum samples collected from 3 hospitals. Figure B shows the spatial distribution of serum samples collected excluding samples from Children’s Hospital 2. "

spatial_3bv <- sero_3bv %>% 
  group_by(district) %>% 
  count() %>% 
  left_join(qhtp, ., by = join_by(varname_2 == district)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = n,geometry = geom),
          show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Number of \nserum samples")+
  theme_void()+
  labs(tag = "A")+
  theme(legend.position = "bottom")


spatial_3bv_woch2 <- sero_3bv %>% 
  filter(hospital != "Bv Nhi Dong 2") %>% 
  group_by(district) %>% 
  count() %>% 
  left_join(qhtp, ., by = join_by(varname_2 == district)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_sf(aes(fill = n,geometry = geom),
          show.legend = T)+
  paletteer::scale_fill_paletteer_c("ggthemes::Classic Red",
                                    na.value="white",
                                    name = "Number of \nserum samples")+
  theme_void()+
  labs(tag = "B")+
  theme(legend.position = "bottom")

spatial_3bv | spatial_3bv_woch2
```

# Hospital-specific age-dependent susceptibility profiles and corresponding hospitalizations

```{r}
hcm_pop_19_age <- census2019 %>% mutate(
  district = district %>% 
    str_replace_all(
      c("Quận 2" = "Thủ Đức",
        "Quận 9" = "Thủ Đức")) %>% 
    str_remove("Quận|Huyện") %>%
    trimws(which = "both") %>% 
    stri_trans_general("latin-ascii") %>% 
    tolower()) %>% 
  filter(age2 <= 15) %>% 
  group_by(district,age2) %>% 
  summarise(pop = sum(n)) %>% 
  ungroup()

plot_s_and_case <- function(hos){
  
  selected_hos <- case_when(
    hos == "CH1" ~ "Children's Hospital 1",
    hos == "CH2" ~ "Children's Hospital 2",
    hos == "CCH" ~ "City Children's Hospital"
  )
  
  hos_sero <- case_when(
    hos == "CH1" ~ "CH1",
    hos == "CH2" ~ "Bv Nhi Dong 2",
    hos == "CCH" ~ "Bv Nhi Dong Tp"
  )
  
  numcases_age <- meases_1812_3hos %>% 
    filter(hospital == selected_hos & 
             year(admission) %in% c(2024,2025) &
             month(admission) >= 4) %>% 
    # mutate(adm_week = as.Date(floor_date(admission, "week"))) %>% 
    filter(age_gr != ">15") %>% 
    group_by(admission,age_gr) %>% 
    count() %>% 
    ggplot(aes(x = admission,y=n))+
    geom_col(position = position_dodge())+
    geom_vline(xintercept = as.Date("2024-04-30"))+
    facet_wrap(~ age_gr, ncol = 1) +
    scale_x_date(limits = c(as.Date("2024-04-01"),as.Date("2025-01-01")),
                 breaks = "2 months",
                 date_labels = "%b %Y",
                 name = "Admission date")+
    labs(y = "Cases per day",tag = hos)+
    theme_bw()
  
  age_profile_424 <- sero_3bv %>% 
    filter(hospital == hos_sero) %>% 
    age_profile() %>% 
    mutate(sus = 1-fit,
           lwr_s = 1-lwr,
           upr_s = 1-upr)
  
  sp_424 <- age_profile_424 %>% 
    ggplot(aes(x = age,y = fit))+
    geom_line()+
    geom_ribbon(aes(ymin = lwr,ymax = upr),fill = "blue",alpha = 0.4)+
    scale_y_continuous(limits = c(0,1),
                       name = "Seroprevalence (%)",
                       labels = scales::label_percent())+
    scale_x_continuous(limits = c(0,16),
                       breaks = seq(0,16,by=2),
                       name = "Age (years)")+
    labs(tag = hos)+
    theme_bw()
  
  district_serum <- sero_3bv %>% 
    filter(hospital == hos_sero) %>% 
    pull(district) %>% unique()
  
  num_sep <- hcm_pop_19_age %>% 
    filter(district %in% district_serum) %>% 
    group_by(age2) %>% 
    summarise(pop = sum(pop)) %>% 
    gam(pop~s(age2,k = 15),data = .) %>% 
    predict(list(age2 = seq(0,16,le=512))) %>% 
    as.tibble() %>% 
    mutate(age = seq(0,16,le=512)) %>% 
    left_join(age_profile_424,., by = join_by(age)) %>% 
    mutate(num_s = sus*value) %>% 
    ggplot(aes(x = age,y = value))+
    geom_line()+
    scale_x_continuous(breaks = seq(0,16,by=2))+
    scale_y_continuous(limits = c(0,120000))+
    labs(y = "Number of susceptible", x = "Age (years)")+
    labs(tag = hos)+
    theme_bw()
  
  a_ch1 <- num_sep/
    sp_424
  
  b_ch1 <- numcases_age
  
  (a_ch1 | b_ch1) + 
    plot_layout(widths = c(1, 2))
  
}
```


## CH1

```{r}
#| fig-width: 10
#| fig-height: 6 
#| out-width: "100%"

plot_s_and_case(hos = "CH1")
```


## CH2

```{r}
#| fig-width: 10
#| fig-height: 6 
#| out-width: "100%"

plot_s_and_case(hos = "CH2")
```


## CCH

```{r}
#| fig-width: 10
#| fig-height: 6 
#| out-width: "100%"

plot_s_and_case(hos = "CCH")
```
